\documentclass[12pt]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{paralist}

\theoremstyle{definition}
\newtheorem{question}{Question}

\begin{document}
\begin{center}
\textbf{\Huge
Lesson Plan: Meeting One
}
\end{center}
\vspace{.5in}

The main goal of meeting one is to get students in the frame of mind to think about probability, and \emph{conditional probability} in particular.

\section*{The Monty Hall Problem}

\subsection*{Intro: Game Show!}

Play Game show host \& do the Monty Hall problem. 

\emph{materials needed:} three cups for ``doors.'' two prizes (small candy bars?), showmanship

Run the game twice, two different student ``volunteers.'' 
After the fun, carefully explain the mechanics of the game. 
Also, give a bit of history by talking about \emph{Let's Make a Deal!} and Monty Hall and Marilyn Vos Savant.

\subsection*{Core Activity: Play the game}

Point out that the interesting question is to find the best strategy for maximizing the chance of winning. 
(No guarantees, but find the better chance.) 
The two strategies are \emph{keep} and \emph{switch}. 

\emph{materials needed:} enough ``markers'' of some sort to use as fake prizes.

Split class into teams of two. 
Have them tear up a sheet of paper to make three little screens labeled A, B, and C. 
Students take turns playing host and contestant. 
Each should play the game 10 times with a single strategy. 
One person plays the \emph{keep} strategy ten times. 
Then they switch roles and the second person plays the \emph{switch} strategy ten times. 
Keep a record of wins and losses.

\subsection*{Surprise: Show the simulation}

Ask the class which seems a better strategy. 
Explain that you are doing a lot of trials because that is what probability is really about: \emph{long run averages}.

Show them simple simulations. 
Explain the code and run it in front of them. 
Discuss.

Sage code for \emph{keep}.

\begin{verbatim}
def stay_check(k=1000):
    doors = [1,2,3]
    win_count , loss_count, trial = 0, 0, 0
    while trial < k:
        winner = randint(1,3)
        pick = randint(1,3)
        for door in doors:
            if door != winner and door != pick:
                open = door
                break
        if pick == winner:
            win_count = win_count + 1
        else:
            loss_count = loss_count + 1
        trial = trial + 1
    print "%d wins and %d losses" % (win_count, loss_count)

stay_check(10^6)
\end{verbatim}

\clearpage

Sage code for \emph{switch}.

\begin{verbatim}
def switch_check(k=1000):
    doors = [1,2,3]
    win_count , loss_count, trial = 0, 0, 0
    while trial < k:
        winner = randint(1,3)
        pick = randint(1,3)
        for door in doors:
            if door != winner and door != pick:
                open = door
                break
        for door in doors:
            if door != pick and door != open:
                new = door
                break
        if new == winner:
            win_count = win_count + 1
        else:
            loss_count = loss_count + 1
        trial = trial + 1
    print "%d wins and %d losses" % (win_count, loss_count)

switch_check(10^6)
\end{verbatim}




\end{document}