%% Fall 2013 MDM Homework One
\documentclass[12pt,letterpaper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm,headheight=22pt]{geometry}
\usepackage{fancyhdr}
\usepackage{setspace}
\usepackage{lastpage}

\begin{document}

%other parameters
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}
\setlength{\parindent}{0pt}

%header and footer parameters
\pagestyle{fancy}
\lhead{Math 1100 section 03}
\chead{Weekly Homework}
\rhead{Due: Friday, August 30}
\lfoot{} 
\cfoot{\emph{Prof. Hitchman}} 
\rfoot{} 

\begin{center}
{
\Large
\textbf{Written Assignment \#1}
}
\end{center}

Please write a short \emph{Mathematical Autobiography}.
I think a single page should be enough.\\

I want to know something about your attitude towards mathematics, and your relationship with the subject.
Feel free to address whatever you want to share.
In case you are feeling at a loss for where to start, consider the following questions:
\begin{itemize}
\item Do you enjoy mathematics? Why or why not?
\item Have you had a particularly good experience studying mathematics?
If so, what made it work for you?
\item Have you had a particularly negative experience with mathematics?
If so, what made it so negative?
\item What do you hope to get out of our class this term?
\end{itemize}
I don't expect you to cover each question. Rather, use these as a starting point for your thinking.\\[.25in]


This is a writing assignment, so I expect attention to standard grammar and usage.
I would prefer a typed and printed page.


\end{document}


















