%---Fall 2013 Math 3600 Syllabus

\documentclass[12pt]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{url}

\begin{document}

\title{Mathematics in Decision Making\\ Math 1100 section 03\\ Fall 2013}

\maketitle

\thispagestyle{empty}

\noindent

\noindent
\textbf{Basic Information:}\\
\begin{tabular}{ll}
Instructor:  & Theron J Hitchman \\
Office: & 327 Wright Hall\\
Office Hours: & \textbf{MWF 11am - noon.}  To meet at another time, make an appointment. \\
contact info: & email:  \url{theron.hitchman@uni.edu} \phantom{spc}  phone: 319-273-2646 \\
Course Web Page: & \url{https://sites.google.com/a/uni.edu/mdm-hitchman/home}  \\
 \phantom{wide} & Check this regularly.
 Lots of important things are at this page. \\[.1in]
\end{tabular}




\noindent
\textbf{Basic Course Policies and Procedures:}\\
This course will have three main units, with an exam after each one.
\begin{itemize}
\item Topology: Knots, tangles and surfaces
\item Infinity: comparing large sets
\item Data Science: some probability and statistics
\end{itemize}

We will have regular homework, too. It will not be “graded,” but it will receive comments to help you improve. Your final grade for this course will be determined as follows:
\begin{itemize}
\item 90\% - Three Unit Exams (30\% each)
\item 10\% - Homework
\end{itemize}
Exam dates will be announced in class, but should roughly break the semester into thirds.\\

There will be two types of assignments for this course:
\begin{itemize}
\item \textbf{Daily Reading and Guided Practice:} To be completed before each class meeting. This is where most of your time should be spent on this course.
\item \textbf{Weekly Assignments:} Once a week I will collect an assigment for commentary and grading.\\[.1in]
\end{itemize}

\noindent
\textbf{Class Materials}\\
There is no text for this course, though I will hand out resources often in class, and post readings and assignments to the course web page. You should expect to have a reading assignment and homework before each class. I will not track attendance officially, but it is important to come to class and be involved.

Occasionally, it will be helpful to have your own manipulatives. These should be easy to acquire or make, and I’ll let you know when the time comes.

We will often use a classroom response system in class. I have chosen Poll Everywhere for this because it allows students to participate using any web-enabled device, or by text messaging. Please bring a smartphone, cell phone, laptop, or tablet to class so you can participate.\\

\clearpage

\noindent
\textbf{Official and Important Stuff You Should Know About:}\\

\noindent
\textbf{Academic Accommodations:}\\
Please address any special needs or special accommodations with me at the beginning of the semester or as soon as you become aware of your needs. Those seeking accommodations based on disabilities should obtain a Student Academic Accommodation Request form from Student Disability Services (SDS) (phone 319-273-2677). SDS is located on the top floor of the Student Health Center, Room 103.

\noindent
\textbf{Academic Learning Center:}\\
I encourage you to utilize the Academic Learning Center’s assistance with writing, math, science, reading, and learning strategies. There is no charge for currently-enrolled UNI students. UNI’s Academic Learning Center is located in 007/008 ITTC. Visit the website at \url{http://www.uni.edu/unialc/} or phone 319-273-2361 for more information.\\


\noindent
\textbf{Equal Opportunity \& Non-Discrimination Statement}\\
\noindent
Purpose:\\
To provide guidelines regarding equal opportunity and non-discrimination at the University in compliance with applicable federal and state non-discrimination and affirmative action laws and regulations.\\

\noindent
Policy Statement:\\
No person shall be excluded from participation in, be denied the benefits of, or be subjected to discrimination in employment, any educational program, or any activity of the University, on the basis of age, color, creed, disability, gender identity, national origin, race, religion, sex, sexual orientation, veteran status, or on any other basis protected by federal and/or state law.\\

The University of Northern Iowa prohibits discrimination and promotes affirmative action in its educational and employment policies and practices as required by Title IX of the Educational Amendments of 1972, the Americans with Disabilities Act of 1990, Section 504 of the Rehabilitation Act of 1973, Title VII of the Civil Rights Act of 1964 and other applicable laws and University policies. The University of Northern Iowa prohibits sexual harassment, including sexual violence.\\

The following person has been designated to handle inquiries regarding the non-discrimination policies and serves as the University Title IX Officer: Leah Gutknecht, Assistant to the President for Compliance and Equity Management, Office of Compliance and Equity Management, 117 Gilchrist Hall, UNI, Cedar Falls, IA 50614-0028, 319-273-2846, \url{leah.gutknecht@uni.edu}

\end{document}





















%sagemathcloud={"zoom_width":100}